---
output:
  html_document: default
  pdf_document: default
  word_document: default
---

---
title: "Prediction field of LMA"
author: "Yunke Peng"
date: "Nov 11 2020"
output: html_document
---

## Introduction about Global nc files

Here LMA was done separately, because (1) pre-processing of TRY maps takes a long time, whcih requires the step to resample data from TRY map to 0.5 grids, (2) knn method also takes a long time, and it necessarily needs global predictors from PPFD, Tg and alpha, therefore this file is better to be done separately.

A brief information about LMA, and three main predictors that will be used in knn.
 * alpha (SPLASH: asssited by David Sandoval)
 * PPFD (umol/m2/s) (WFDEI: http://www.eu-watch.org/gfx_content/documents/README-WFDEI%20(v2016).pdf)
 * Tg (degree celcius) (CRU ts 4.01: https://crudata.uea.ac.uk/cru/data/hrg/cru_ts_4.01/)
 * LMA (g/m2) (TRY Global traits map 3km: https://isp.uv.es/code/try.html)

alpha, PPFD, Tg will directly use 0.5 * 0.5 grids data from final_ncfile/

1. Pre-processing of LMA, including resample TRY traits map.
```{r}

library(rbeni)
library(raster)

#1. Input 0.5 resolution df, and convert to raster before resampling (e.g. elevation.nc)
elev_nc <- read_nc_onefile("~/data/watch_wfdei/WFDEI-elevation.nc")
elev <- as.data.frame(nc_to_df(elev_nc, varnam = "elevation"))
summary(elev)
names(elev) <- c("lon","lat","z")
coordinates(elev) <- ~lon+lat 
gridded(elev) <- TRUE
raster_z <- raster(elev, "z") 

bounding_box <- extent(-180, 180, -90, 90)
raster_z_crop <- crop(raster_z, bounding_box)
raster_z_crop

#2. Input 3km LMA raster that needs resampled
#Original data source: https://isp.uv.es/code/try.html
raster_SLA <- raster("~/data/TRY_maps/data_orig/SLA_3km_v1.tif")
raster_SLA_crop <- crop(raster_SLA, bounding_box)
raster_SLA_crop

#raster of SLA (0.02694946* 0.02694946) will be resampled to 0.5*0.5 resolution, based on provided elevation raster 
res(raster_z_crop)
res(raster_SLA_crop)

resampled_SLA <- resample(raster_SLA_crop, raster_z_crop, method="ngb")
df_SLA <- stack(resampled_SLA)
df_SLA <- as.data.frame(df_SLA,xy = TRUE)
names(df_SLA) <- c("lon","lat","SLA")
summary(df_SLA)

df_SLA2 <- subset(df_SLA,(SLA > 0 & SLA < 1000)) # remove some amazingly wrong number (negative or too high!)
df_SLA2$LMA <- 1000/df_SLA2$SLA #mm2/mg -> g/m2
LMA_df <- df_SLA2[,c("lon","lat","LMA")]  

``` 


2. knn method, using Tg, PPFD, alpha and lat to fill empty areas in LMA map, as shown above
```{r}
library(raster)
library(ncdf4)
library(dplyr)      # for data wrangling
library(ggplot2)    # for awesome graphics
library(rsample)    # for creating validation splits
library(recipes)    # for feature engineering
# Modeling packages
library(caret)       # for fitting KNN models
library(h2o)       # for resampling and model training
library(AmesHousing)
library(modeldata)
library(rsample)
library(dslabs)
library(purrr)
library(randomForest)

# recap
summary(LMA_df)
dim(LMA_df)

#Map after resamping
plot_map3(LMA_df, 
          varnam = "LMA",plot_title = " LMA (g m-2)",
          latmin = -65, latmax = 85)

#We start our work from here - using knn method to fill our grids

#(1) first, input Tg, PPFD, alpha directly from final_nc/ (the data directly used in nimpl simulation)
# NA value was converted to 9999 in previous step (to prevent FPE in nimpl), therefore, converting it back
Tg_df <- nc_to_df(read_nc_onefile("~/data/Final_ncfile/Tg.nc"), varnam = "Tg")
Tg_df$myvar[Tg_df$myvar == 9999] <- NA
names(Tg_df) <- c("lon","lat","Tg")

PPFD_df <- nc_to_df(read_nc_onefile("~/data/Final_ncfile/PPFD.nc"), varnam = "PPFD")
PPFD_df$myvar[PPFD_df$myvar == 9999] <- NA
names(PPFD_df) <- c("lon","lat","PPFD")

alpha_df <- nc_to_df(read_nc_onefile("~/data/Final_ncfile/alpha.nc"), varnam = "alpha")
alpha_df$myvar[alpha_df$myvar == 9999] <- NA
names(alpha_df) <- c("lon","lat","alpha")

merge1 <-Reduce(function(x,y) merge(x = x, y = y, by = c("lon","lat"),all.x=TRUE), 
                list(Tg_df,PPFD_df,alpha_df,LMA_df))
summary(merge1) # we will start knn from here, it is clear that LMA has less data, and this is what we aim to interpolate based on knn

merge2 <- subset(merge1,PPFD>0) # only select available grids (in land). This will be furtherly used to determine training and testing (predicted) data.
dim(merge2)

#(2) start knn

all_data <- merge2[,c("LMA","lat","Tg","alpha","PPFD")] # this is final input will be used in knn, which removed lon here.

training_data <- subset(all_data,LMA>0) #only select available LMA data as training data

testing_data <- subset(all_data,is.na(LMA)==TRUE)  #only select NA-LMA data as testing data, which will be used in knn at the end.

dim(training_data)
dim(testing_data)

# we will use those 19020 grids to predict those 44816 grids

# 3. knn training
cv <- trainControl(
  method = "cv", 
  number = 10
)

# Create a hyperparameter grid search
hyper_grid <- expand.grid(
  k = floor(seq(5, 20, length.out = 15))
)

knn_fit <- train(LMA ~., data = training_data,
                 method = "knn",
                 trControl=cv,
                 preProcess = c("center", "scale"),
                 tuneGrid = hyper_grid)

knn_fit # it indicated k =9 as the best design, with R2=0.61, RMSE = 10.58

# output prediction data and add it in testing_data
test_pred <- predict(knn_fit,newdata=testing_data)
testing_data$LMA <- test_pred

#rbind, and merge back with lon.
final <- rbind(training_data,testing_data)

lon_data <- merge2[,c("lon","lat","Tg","alpha","PPFD")]

final2 <-Reduce(function(x,y) merge(x = x, y = y, by = c("lat","Tg","alpha","PPFD"),all.x=TRUE), 
                list(final,lon_data))

#form final3 dataframe as final output, and plot
LMA_knn_df <- final2[,c("lon","lat","LMA")]

#This map is LMA after knn method
plot_map3(LMA_knn_df, 
          varnam = "LMA",plot_title = " LMA (g m-2)",
          latmin = -65, latmax = 85)

#Finally, ready to be outputted! It merges from 63836 grids to 259200 cells, which makes it as a perfect standard for nimpl input
LMA_knn_df2 <-Reduce(function(x,y) merge(x = x, y = y, by = c("lon","lat"),all.x=TRUE), 
                list(merge1[,c("lon","lat")],LMA_knn_df))

summary(LMA_knn_df2)
dim(LMA_knn_df2)

#And of course, convert NA to 9999 before nimpl, then apply cdo process.
#LMA_knn_df2[is.na(LMA_knn_df2)] <- 9999
#coordinates(LMA_knn_df2) <- ~lon+lat 
#gridded(LMA_knn_df2) <- TRUE
#rLMA <- raster(LMA_knn_df2, "LMA") 

#d <- writeRaster(rLMA, paste0("~/data/nimpl_sofun_inputs/","LMA_1107.nc"),format = "CDF", overwrite = TRUE, varname = "LMA", varunit = "g/m2", longname = "Leaf-mass-per-area", xname = "lon", yname = "lat")

``` 

3. test map's LMA data comparing with obs. LMA data
```{r}

#3. This is not the end! Let's check how LMA from map closed to obs. LMA
#We will try both version, one is LMA from orginal data (before knn), another is final data (after knn and exactly the one used for )
dim(LMA_df) # org. data has less grids, as we all know!
dim(LMA_knn_df)

library(spgwr)

#Now, input a large global dataset of LMA individuals data
SP_input <- read.csv(file="~/data/CN_leaf/final_individuals.csv") #all individuals
SP_input2 <- SP_input[,c("lat","lon","Elevation","lma")]
sitemean <- aggregate(SP_input2,by=list(SP_input2$lon,SP_input2$lat), FUN=mean, na.rm=TRUE) #site-mean
leafcn_site <- sitemean[,c("lon","lat","Elevation","lma")]
names(leafcn_site) <- c("lon","lat","z","lma")


#start interpolating sites
###using direct method - based on original data (before knn)
leafcn_global <- na.omit(LMA_df)
coordinates(leafcn_global) <- ~lon+lat 
gridded(leafcn_global) <- TRUE
rleafcn_global <- raster(leafcn_global, "LMA") 
sp_sites <- SpatialPoints(leafcn_site[,c("lon","lat","z")]) # only select lon and lat

leafcn_pred_direct <- extract(rleafcn_global, sp_sites, sp = TRUE) %>% as_tibble() %>% 
  right_join(leafcn_site, by = c("lon", "lat","z")) %>% 
  dplyr::rename( TRY_lma = LMA)

leafcn_pred_direct <- as.data.frame(leafcn_pred_direct)
head(leafcn_pred_direct)

#using direct method
analyse_modobs2(leafcn_pred_direct,"TRY_lma","lma", type = "points")


###using direct method - based on final data (after knn)
leafcn_global <- na.omit(LMA_knn_df)
coordinates(leafcn_global) <- ~lon+lat 
gridded(leafcn_global) <- TRUE
rleafcn_global <- raster(leafcn_global, "LMA") 
sp_sites <- SpatialPoints(leafcn_site[,c("lon","lat","z")]) # only select lon and lat

leafcn_pred_direct <- extract(rleafcn_global, sp_sites, sp = TRUE) %>% as_tibble() %>% 
  right_join(leafcn_site, by = c("lon", "lat","z")) %>% 
  dplyr::rename( TRY_lma = LMA)

leafcn_pred_direct <- as.data.frame(leafcn_pred_direct)
head(leafcn_pred_direct)

#using direct method
analyse_modobs2(leafcn_pred_direct,"TRY_lma","lma", type = "points")

#The comparsion between map's extracterd site LMA and obs. LMA is amazingly poor! That's probably why we fit poor performance in leaf c/n prediciton in nimpl simulation


```


4. Add site PPFD and Tg into model, to see if it helps to predict better
```{r}

dim(leafcn_pred_direct)
merge3 <- merge2[,c("lon","lat","Tg","PPFD","alpha")]
elev_nc <- read_nc_onefile("~/data/watch_wfdei/WFDEI-elevation.nc")
elev <- as.data.frame(nc_to_df(elev_nc, varnam = "elevation"))
names(elev) <- c("lon","lat","z")

all_global <-as.data.frame(Reduce(function(x,y) merge(x = x, y = y, by = c("lon","lat"),all.x=TRUE), 
                list(elev,merge3)))

leafcn_pred_direct <- leafcn_pred_direct[,c("lon","lat","z","lma","TRY_lma")]

a <- 1.5

#interpolating Tg and PPFD
for (i in c(1:nrow(leafcn_pred_direct))){
  gpp_global <- na.omit(all_global)
  gpp_part <- subset(gpp_global,lon>(leafcn_pred_direct[i,1]-a)&lon<(leafcn_pred_direct[i,1]+a)&
                       lat>(leafcn_pred_direct[i,2]-a)&lat<(leafcn_pred_direct[i,2]+a))
  coordinates(gpp_part) <- c("lon","lat")
  gridded(gpp_part) <- TRUE
  
  gpp_coord <- leafcn_pred_direct[i,1:3]
  coordinates(gpp_coord) <- c("lon","lat")
  leafcn_pred_direct$Tg[i] <- (gwr(Tg ~ z, gpp_part, bandwidth = 1.06, fit.points =gpp_coord,predictions=TRUE))$SDF$pred
  leafcn_pred_direct$PPFD[i] <- (gwr(PPFD ~ z, gpp_part, bandwidth = 1.06, fit.points =gpp_coord,predictions=TRUE))$SDF$pred
}

#newly include narea and vcmax25

SP_input <- read.csv(file="~/data/CN_leaf/final_individuals.csv") #all individuals
SP_input2 <- SP_input[,c("lat","lon","Elevation","narea","Vcmax.25")]
sitemean2 <- aggregate(SP_input2,by=list(SP_input2$lon,SP_input2$lat), FUN=mean, na.rm=TRUE) #site-mean
sitemean2 <- sitemean2[,c("lon","lat","Elevation","narea","Vcmax.25")]
names(sitemean2) <- c("lon","lat","z","narea","Vcmax.25")

final <-as.data.frame(Reduce(function(x,y) merge(x = x, y = y, by = c("lon","lat","z"),all.x=TRUE), 
                list(leafcn_pred_direct,sitemean2)))
summary(final)

final$nmass <- final$narea/final$lma

#2 models to be considered now
summary(lm(nmass ~ Tg + PPFD + lma , data = final))

final$vcmax25_lma <- final$Vcmax.25/final$lma

summary(lm(nmass ~ vcmax25_lma, data = final))

dim(final)

```