---
title: "Global input for all prediction field"
author: "Yunke Peng"
date: "Nov 18 2020"
output: html_document
---

## Introduction about Global nc files

Elements below are global inputs that will be used in statistical model-based calculations. All data has a final resolution as 0.5 * 0.5 degree - though the resample procedure was needed for soil C/N and LMA. All climate or environmental data (fAPAR, alpha, PPFD, Tg, vpd) has constrained the year from 1982 to 2011 (30 yrs average); while some others (soil C/N, age and LMA) also had a long-term average, according to their paper's description.

 * Soil C/N (ISRIC; previously asssited by Nick Smith and Trevor Keenan in 2019)
 * stand age (yrs) (Poulter et al. 2019: https://doi.pangaea.de/10.1594/PANGAEA.889943)
 * fAPAR (3g GIMMS: https://developers.google.com/earth-engine/datasets/catalog/NASA_GIMMS_3GV0)
 * alpha (SPLASH: asssited by David Sandoval)
 * PPFD (umol/m2/s) (WFDEI: http://www.eu-watch.org/gfx_content/documents/README-WFDEI%20(v2016).pdf)
 * Tg (degree celcius) (CRU ts 4.01: https://crudata.uea.ac.uk/cru/data/hrg/cru_ts_4.01/)
 * vpd (kPa) (CRU ts 4.01: https://crudata.uea.ac.uk/cru/data/hrg/cru_ts_4.01/)
 * LMA (g/m2) (TRY Global traits map 3km: https://isp.uv.es/code/try.html)
 * Vcmax25 (umol/m2/s) (will be further processed in Beni Stocker's SOFUN in Euler).
 
The process below will provide original codes of soil C/N, stand age, fAPAR, alpha and LMA, as well as description text of PPFD, Tg, vpd. Vcmax25 will be seperately processed in SOFUN in Euler then.


#1. soil C/N

```{r}
library(raster)
library(rgdal)
library(dplyr)
library(rbeni)
library(ncdf4)

#see soilCN_field.rmd

######## the code below is old method --> not used now.

## get resolution of original file
#res_raster <- res(raster_soil)

## aggregate to 0.5 deg
#res_out <- c(0.5, 0.5)
#agg <- res_out / res_raster
#raster_soil_agg <- raster::aggregate(raster_soil, fact = agg, fun = mean, na.rm = TRUE, expand = FALSE)

#writeRaster(raster_soil_agg,"~/data/ISRIC/data_halfdeg/cn_soil_isric_halfdeg.nc",overwrite=TRUE, format="CDF", varname="cnsoil", varunit="unit", longname="Soil C:N ratio", xname="lon", yname="lat")

## make sure you get an array of 720 x 360
#nc <- read_nc_onefile("~/data/ISRIC/data_halfdeg/cn_soil_isric_halfdeg.nc")


#soil <- raster('E:/final round/data/w001000.adf') # map input from the last 2015 version of ISRIC (Batjes 2015 ISRIC report)
# 
# Global_grid <- as.data.frame(read.csv(file="~/data/ISRIC/data_orig/data/Global_cell.csv")) 
# #Global_grid <- as.data.frame(read.csv(file="E:/climate/Global_cell.csv")) #Input global land grids (0.5 degree;N = 259200)
# 
# Global_grid$ID <- c(1:259200) #create a list of number, just helping to form a raster then.
# 
# pointXY<-SpatialPointsDataFrame(Global_grid[,c('lon','lat')], data=data.frame(unique.id=paste(Global_grid$ID))) # convert to spatial data format 
# 
# vec.soil.extract <-extract(soil,pointXY) 
# data.soil.extract <-data.frame(pointXY,SUID=vec.soil.extract) 
# # These two steps are important, it will extract SUID information from soil map to our grids data, so that each our global grid will have a SUID, which will be further combined (merged) with soil C/N data then.
# 
# ISRIC.data<-read.csv(file="~/data/ISRIC/data_orig/data/HW30s_FULL.csv",header=TRUE,sep=";",dec = ".") # Now, input ISRIC database
# 
# data.soil.extract <- merge(data.soil.extract,ISRIC.data,by='SUID',all.x=TRUE) # merge site with soil variables by using SUID
# 
# data.soil.extract2 <- subset(data.soil.extract,CNrt>0) # select available CNrt
# data.soil.extract3 <- data.soil.extract2[,c("lon","lat","CNrt")] # only select CNrt variable
# 
# ss1 <- aggregate(data.soil.extract3,by=list(data.soil.extract3$lon,data.soil.extract3$lat), FUN=mean, na.rm=TRUE) # note that in each grid there might be more than 1 samples measured, so we should aggregate them which make sures that one grid holds one data only.
# 
# ss2 <- ss1[,c(3,4,5)] # now,select lon, lat and CNrt only
# 
# final_CN <-Reduce(function(x,y) merge(x = x, y = y, by = c("lon","lat"),all.x=TRUE), 
#                    list(Global_grid,ss2)) #merged this with global grids (N = 259200)
# 
# df_CN <- final_CN[,c("lon","lat","CNrt")]
# 
# #convert dataframe to raster
# coordinates(final_CN) <- ~lon+lat 
# gridded(final_CN) <- TRUE
# rCNrt <- raster(final_CN, "CNrt") 
# plot(rCNrt)
# 
#prepare lon and lat
ncin <- nc_open("~/data/watch_wfdei/WFDEI-elevation.nc")
lon <- ncvar_get(ncin,"lon")
lat<-ncvar_get(ncin,"lat")

# #output nc file - In Euler its path is same: "~/data/nimpl_sofun_inputs/map/Final_ncfile"
# summary(df_CN)
# CNrt <- list(df_to_grid(df_CN,varnam = "CNrt", lonnam = "lon", latnam = "lat"))
# names(CNrt) <- "CNrt"
# varams = "CNrt"
# test <- list(lon,lat,CNrt,varams)
# names(test) <- c("lon","lat","vars","varams")
# write_nc2(test,varnams = "CNrt",long_name = "Soil carbon to nitrogen ratio",units = "unitness",
#           path = "~/data/nimpl_sofun_inputs/map/Final_ncfile/CNrt.nc")

```


#2. stand age
```{r}
library(raster)
library(ncdf4)
library(dplyr)
library(maps)

#set the path and filename
ncfname <- paste ("~/data/GFAD/data_orig/", "GFAD_V1-1", ".nc", sep="")
#ncfname <- paste ("E:/C-N cycling/Carbon allocation/672 sites analysis/stand age/GFAD2/", "GFAD_V1-1", ".nc", sep="")

#1. open a netCDF file and check variable
ncin <- nc_open(ncfname)

#2. Get coordinate and value
lon <- ncvar_get(ncin,"lon")
nlon <- dim(lon) 
lat<-ncvar_get(ncin,"lat")
nlat<-dim(lat)
age <-ncvar_get(ncin,"age")
nc_close(ncin)

pre.vec.long <- as.vector(age)
pre.mat <- matrix(pre.vec.long, nrow = nlon * nlat, ncol = 4*15) # 259200 columns * 60 rows - here 60 represents 4 PFTs and 15 age classes.

lonlat <- expand.grid(lon, lat)
lonlat$age <- rowSums(pre.mat[,1:4])*5+rowSums(pre.mat[,5:8])*15+rowSums(pre.mat[,9:12])*25+rowSums(pre.mat[,13:16])*35+rowSums(pre.mat[,17:20])*45+rowSums(pre.mat[,21:24])*55+rowSums(pre.mat[,25:28])*65+rowSums(pre.mat[,29:32])*75+rowSums(pre.mat[,33:36])*85+rowSums(pre.mat[,37:40])*95+rowSums(pre.mat[,41:44])*105+rowSums(pre.mat[,45:48])*115+rowSums(pre.mat[,49:52])*125+rowSums(pre.mat[,53:56])*135+rowSums(pre.mat[,57:60])*145
names(lonlat) <- c("lon","lat","age")

age_input <- as.data.frame(lonlat)
summary(age_input)
dim(age_input)

coordinates(age_input) <- ~lon+lat 
gridded(age_input) <- TRUE
r <- raster(age_input, "age") 
plot(r) # now it shows an empty grid in lower AUS, also in some other regions. We need to interpolate them based on local mean value.

#2. Input continent map
ncfname <- paste ("~/data/landmasks/", "srex-region-masks_20120709.srex_mask_SREX_masks_all.05deg.time-invariant", ".nc", sep="")
#ncfname <- paste ("D:/PhD/nimpl_sofun_inputs/Data/srex/", "srex-region-masks_20120709.srex_mask_SREX_masks_all.05deg.time-invariant", ".nc", sep="")
ncin <- nc_open(ncfname)

lon <- ncvar_get(ncin,"lon")
nlon <- dim(lon) 
lat<-ncvar_get(ncin,"lat")
nlat <- dim(lat)
area <- ncvar_get(ncin,"srex_mask")
nc_close(ncin)

area1 <- as.vector(area)

lonlat <- expand.grid(lon, lat)
continent_input <- as.data.frame(cbind(lonlat,area1))
names(continent_input) <- c("lon","lat","area")

#merge continents and age
continent_age <- merge(continent_input, age_input, by=c("lon","lat"), all.x = T, sort=F )
summary(continent_age)

#3. Input fAPAR, and merge it with continent_age, and convert fAPAR's NA grid to those age = NA
ncfname <- paste ("~/data/fAPAR/fAPAR3g_v2/", "fAPAR3g_v2_1982_2016_FILLED", ".nc", sep="")
#ncfname <- paste ("D:/PhD/nimpl_sofun_inputs/Data/fAPAR/", "fAPAR3g_v2_1982_2016_FILLED", ".nc", sep="")
dname <- "FAPAR_FILLED"

ncin <- nc_open(ncfname)

lon <- ncvar_get(ncin,"LON")
nlon <- dim(lon) 

lat<-ncvar_get(ncin,"LAT")
nlat <- dim(lat)

FAPAR <- ncvar_get(ncin,"FAPAR_FILLED")
nc_close(ncin)

pre.vec.long <- as.vector(FAPAR)

pre.mat <- matrix(pre.vec.long, nrow = nlon * nlat, ncol = 420) # 259200 * 420 (420 is 35 years: 1982-2016)

fAPAR1982_2011 <- pre.mat[,c(1:360)] #extrat fAPAR from 1982-2011

final_fAPAR <- rowMeans(fAPAR1982_2011,na.rm = TRUE) # calculate means

lonlat <- expand.grid(lon, lat)
fAPAR_input <- as.data.frame(cbind(lonlat,final_fAPAR))
names(fAPAR_input) <- c("lon","lat","fAPAR")
summary(fAPAR_input)

#here we merge fAPAR with age and continent data.
continent_age_fAPAR <- merge(continent_age, fAPAR_input, by=c("lon","lat"), all.x = T, sort=F )
summary(continent_age_fAPAR)
dim(continent_age_fAPAR)

#when fAPAR = NA, then those grid's age also = NA. After this, age = 0 only had one possibility, that is, missing data. We will subset them later on and replace them to mean value of local continent.
for (i in 1:nrow(continent_age_fAPAR)){
  if (is.na(continent_age_fAPAR$fAPAR[i]) == TRUE){
    continent_age_fAPAR$new_age[i]<- NA} else { 
      continent_age_fAPAR$new_age[i]<- continent_age_fAPAR$age[i]} 
}

continent_age_fAPAR2 <- continent_age_fAPAR[,c("lon","lat","new_age")]

coordinates(continent_age_fAPAR2) <- ~lon+lat 
gridded(continent_age_fAPAR2) <- TRUE
r <- raster(continent_age_fAPAR2, "new_age") 
plot(r) #we will start working from this map! Because it actually clarfies which grid stand-age = NA (in white), which grid stand-age has missed data (which temporaily considered = 0, and shown in grey, will be interpolated soon) .


#4. Now, start interpolating each continent's missing data (when age = 0) as mean value of local continent
age_final <- as.data.frame(continent_age_fAPAR[,c("lon","lat","area","new_age")])

mylist <- vector(mode = "list", length =25) # for specifying gridded data at certain degrees

for (i in c(1:8,10:24)){ # area 9 and 25 did not have such problem (missing data of stand-age; i.e. age = 0)
  empty_age <- subset(age_final, (area == i & new_age == 0)) #find missing grids (new_age ==0, classfied as missing data)
  available_age <- subset(age_final, (area == i & new_age > 0)) #find non-missing grids and calculate mean
  empty_age$final_age <- mean(available_age$new_age)
  mylist[[i]] <- empty_age
  #print(i)
}

all_age <- as.data.frame(do.call(rbind, mylist))
dim(all_age)

#South AUS (area = 26), however, was interpolated by area =25 's mean stand-age value

South_AUS_empty_age <- subset(age_final, (area == 26 & new_age == 0) )
North_AUS_available_age <- subset(age_final, (area == 25 & new_age > 0))
South_AUS_empty_age$final_age <- mean(North_AUS_available_age$new_age)

all_age2 <- rbind(all_age,South_AUS_empty_age)

all_age3 <- all_age2[,c("lon","lat","final_age")]

#5. merge new interpolated subset of age dataframe into last complete dataframe, and generating findal data to be used.
final <- merge(continent_age_fAPAR, all_age3, by=c("lon","lat"), all.x = T,sort = T)

for (i in 1:nrow(final)){
  if (is.na(final$final_age[i]) == TRUE){
    final$age_used[i] <- final$new_age[i]} else {
      final$age_used[i] <- final$final_age[i]} 
}

final2 <- final[,c("lon","lat","age_used")]
names(final2) <- c("lon","lat","age")

final_data <- final2

coordinates(final2) <- ~lon+lat 
gridded(final2) <- TRUE
r <- raster(final2, "age") 
#here is our final map!
plot(r) 

#6. convert dataframe to nc file as input in nimpl project
summary(final_data)
outlier <- (subset(final_data,age ==0))
dim(outlier)
#map("world", fill=TRUE, col="white", bg="white", ylim=c(-90, 90),xlim = c(-180,180), mar=c(1,0,0,1))
#points(outlier$lon,outlier$lat, col="red", pch=16,cex=1.5)

# it seems to still show 407 grids = 0 (normally in edge or sea). They may not be covered by fAPAR and continent grids from previous analysis, let's just simply assume them as NA.

final_data$age[final_data$age == 0] <- NA

summary(final_data)
#now final_data is ready to be used, let's firstly change its lat, lon order correctly, then convert all NA to 9999

final_data2 <- final_data[order(final_data[,2],final_data[,1]),]

df_age <- final_data2

#output age map
coordinates(final_data2) <- ~lon+lat 
gridded(final_data2) <- TRUE
rage <- raster(final_data2, "age") 
plot(rage)

#output nc file - In Euler its path is same: "~/data/nimpl_sofun_inputs/map/Final_ncfile"
summary(df_age)
age_nc <- list(df_to_grid(df_age,varnam = "age", lonnam = "lon", latnam = "lat"))
names(age_nc) <- "age"
varams = "age"
test <- list(lon,lat,age_nc,varams)
names(test) <- c("lon","lat","vars","varams")
write_nc2(test,varnams = "age",long_name = "stand age",units = "years",
          path = "~/data/nimpl_sofun_inputs/map/Final_ncfile/age.nc")
``` 

#3. fAPAR

```{r}
library(ncdf4)
library(raster)

ncfname <- paste ("~/data/fAPAR/fAPAR3g_v2/", "fAPAR3g_v2_1982_2016_FILLED", ".nc", sep="")

#ncfname <- paste ("E:/C-N cycling/Carbon allocation/fAPAR/fAPAR HUANYUAN/", "fAPAR3g_v2_1982_2016_FILLED", ".nc", sep="")
dname <- "FAPAR_FILLED"

ncin <- nc_open(ncfname)

lon <- ncvar_get(ncin,"LON")
nlon <- dim(lon) 

lat<-ncvar_get(ncin,"LAT")
nlat <- dim(lat)

FAPAR <- ncvar_get(ncin,"FAPAR_FILLED")
nc_close(ncin)

pre.vec.long <- as.vector(FAPAR)

pre.mat <- matrix(pre.vec.long, nrow = nlon * nlat, ncol = 420) # 259200 * 420 (420 is 35 years: 1982-2016)

fAPAR1982_2011 <- pre.mat[,c(1:360)] #extrat fAPAR from 1982-2011

final_fAPAR <- rowMeans(fAPAR1982_2011,na.rm = TRUE)

lonlat <- expand.grid(lon, lat)
fAPAR_input <- as.data.frame(cbind(lonlat,final_fAPAR))
names(fAPAR_input) <- c("lon","lat","fAPAR")

df_fAPAR <- fAPAR_input

coordinates(fAPAR_input) <- ~lon+lat 
gridded(fAPAR_input) <- TRUE
r3 <- raster(fAPAR_input, "fAPAR") 
plot(r3)

#output nc file - In Euler its path is same: "~/data/nimpl_sofun_inputs/map/Final_ncfile"
df_fAPAR$fAPAR[df_fAPAR$fAPAR == "NaN"] <- NA
summary(df_fAPAR)

fAPAR_nc <- list(df_to_grid(df_fAPAR,varnam = "fAPAR", lonnam = "lon", latnam = "lat"))
names(fAPAR_nc) <- "fAPAR"
varams = "fAPAR"
test <- list(lon,lat,fAPAR_nc,varams)
names(test) <- c("lon","lat","vars","varams")
write_nc2(test,varnams = "fAPAR",long_name = "fAPAR",units = "unitless",
          path = "~/data/nimpl_sofun_inputs/map/Final_ncfile/fAPAR.nc")
```

#4. alpha

```{r}
library(raster)
alphalist <- list.files(path = "~/data/alpha/data_orig/",full.names = T) # here this file includes 116 data - that is the annual alpha from 1901 to 2016

#alphalist <- list.files(path = "E:/C-N cycling/Carbon allocation/672 sites analysis/alpha/",full.names = T)

#Input alpha from a list of data frame from 1901 to 2016 (list 116)
empty_alpha <- data.frame(matrix(NA)) 

#calculate alpha within 30 years (1982-2011)
for (i in 82:111){ #here 82:111 means 1982-2011
  load(file = alphalist[i])
  empty_alpha[1:259200,i-81] <- rowMeans(SP_result_monthly,na.rm = TRUE)
}

final_alpha <- rowMeans(empty_alpha,na.rm = TRUE)

alpha_input <- as.data.frame(cbind(lonlat,final_alpha)) 
names(alpha_input) <- c("lon","lat","alpha")

df_alpha <- alpha_input

coordinates(alpha_input) <- ~lon+lat 
gridded(alpha_input) <- TRUE

r4 <- raster(alpha_input, "alpha") 
plot(r4)

#output nc file - In Euler its path is same: "~/data/nimpl_sofun_inputs/map/Final_ncfile"
df_alpha$alpha[df_alpha$alpha == "NaN"] <- NA
summary(df_alpha)

alpha_nc <- list(df_to_grid(df_alpha,varnam = "alpha", lonnam = "lon", latnam = "lat"))
names(alpha_nc) <- "alpha"
varams = "alpha"
test <- list(lon,lat,alpha_nc,varams)
names(test) <- c("lon","lat","vars","varams")
write_nc2(test,varnams = "alpha",long_name = "alpha",units = "unitless",
          path = "~/data/nimpl_sofun_inputs/map/Final_ncfile/alpha.nc")

``` 

#5. PPFD (description and main calculation were provided)
PPFD is instantaneous photosynthetic photon flux density (umol/m2/s). We used the 30 yrs average data (1982-2011), which was originally derived from time-integrated (24 hrs) daily Shortwave radiation (SWdown) from WFDEI. In order to convert from SWdown to PPFD, we need to calculate them as:

PPFD (umol/m2/s) = SWdown (w/m2) * 4.6 * 0.5 

where Solar radiation was converted to incident PPFD assuming a flux : energy ratio of 4.6 umol/J and a photosynthetically active fraction of 0.50 (Peng et al. 2020; New Phytologist).

The code below will represent the global map of final output:

```{r}
library(raster)
library(ncdf4)
ncfname <- paste ("~/data/Final_ncfile/old_9999/", "PPFD", ".nc", sep="")
#ncfname <- paste ("E:/R code/CN MS/PhD/NC files/", "PPFD", ".nc", sep="")
dname <- "PPFD"
ncin <- nc_open(ncfname)

lon <- ncvar_get(ncin,"lon")
nlon <- dim(lon) 
lat<-ncvar_get(ncin,"lat")
nlat <- dim(lat)
PPFD <- ncvar_get(ncin,"PPFD")

nc_close(ncin)

pre.vec.long <- as.vector(PPFD)
pre.vec.long[pre.vec.long == 9999] <- NA

lonlat <- expand.grid(lon, lat)
PPFD_input <- as.data.frame(cbind(lonlat,pre.vec.long))
names(PPFD_input) <- c("lon","lat","PPFD")

df_PPFD <- PPFD_input

coordinates(PPFD_input) <- ~lon+lat 
gridded(PPFD_input) <- TRUE
rPPFD <- raster(PPFD_input, "PPFD") 
plot(rPPFD)

#output nc file - In Euler its path is same: "~/data/nimpl_sofun_inputs/map/Final_ncfile"
summary(df_PPFD)

PPFD_nc <- list(df_to_grid(df_PPFD,varnam = "PPFD", lonnam = "lon", latnam = "lat"))
names(PPFD_nc) <- "PPFD"
varams = "PPFD"
test <- list(lon,lat,PPFD_nc,varams)
names(test) <- c("lon","lat","vars","varams")
write_nc2(test,varnams = "PPFD",long_name = "PPFD",units = "umol m-2 s-1",
          path = "~/data/nimpl_sofun_inputs/map/Final_ncfile/PPFD.nc")

``` 

#6. Tg (description and main calculation were provided)
Tg is growth temperature (degree celcius). We used the 30 yrs average data (1982-2011), which was originally derived from monthly maximum temperature (Tmx) and minimum temperature (Tmn) from CRU ts 4.01. See formula below:

```{r}
#Tg_cal <- function(tmn,tmx){
  #-----------------------------------------------------------------------
  # Input:   sai (s; solar delclination), 12 month constant values
  # Input:   tmn, in monthly or daily degrees
  # Input:   tmx, in monthly or daily degrees
  # Input:   lat, in degrees
  # Output:  Tg,  in monthly or daily degrees
  # Features: Converts growth temperature from tmn, tmx, lon
  #-----------------------------------------------------------------------
#  s1 <- -20.205
#  s2 <- -12.65
#  s3 <- -1.95
#  s4 <- 9.45
#  s5 <- 18.35
#  s6 <- 22.55
#  s7 <- 20.75
#  s8 <- 13.45
#  s9 <- 2.9
#  s10 <- -8.45
#  s11 <- -17.85
#  s12 <- -22.355
#  s <- c(s1,s2,s3,s4,s5,s6,s7,s8,s9,s10,s11,s12)
#  x <-data.frame(matrix(nrow = 259200,ncol=12)) ##create x dataframe: nrow=No. of sites, ncol=monthly step
#  Tg <-data.frame(matrix(nrow = 259200,ncol=12))##see above
#  
#  for (i in 1:12){
#  x[,i]<- -tan(pi*lat/180)*tan(s[i]*pi/180)
#  Tg[,i]<-tmx[,i]*(0.5+(1-x[,i]^2)^(0.5)/(2*acos(x[,i])))+ tmn[,i]*(0.5-(1-x[,i]^2)^(0.5)/(2*acos(x[,i])))
#    }
#  Tg[Tg<0] <-NA
#  Tg[Tg =="NaN"] <- NA
#  return(Tg)
#}

#where Tg was calculated from Tmn and Tmx based on monthly average data (Peng et al. 2020 submitted to #GCB).The code below will represent the global map of final output:

library(raster)
library(ncdf4)
ncfname <- paste ("~/data/Final_ncfile/old_9999/", "Tg", ".nc", sep="")
#ncfname <- paste ("E:/R code/CN MS/PhD/NC files/", "Tg", ".nc", sep="")
dname <- "Tg"

ncin <- nc_open(ncfname)

lon <- ncvar_get(ncin,"lon")
nlon <- dim(lon) 
lat<-ncvar_get(ncin,"lat")
nlat <- dim(lat)

Tg <- ncvar_get(ncin,"Tg")
nc_close(ncin)

pre.vec.long <- as.vector(Tg)
pre.vec.long[pre.vec.long == 9999] <- NA

lonlat <- expand.grid(lon, lat)
Tg_input <- as.data.frame(cbind(lonlat,pre.vec.long))
names(Tg_input) <- c("lon","lat","Tg")

df_Tg <- Tg_input

coordinates(Tg_input) <- ~lon+lat 
gridded(Tg_input) <- TRUE
rTg <- raster(Tg_input, "Tg") 
plot(rTg)

#output nc file - In Euler its path is same: "~/data/nimpl_sofun_inputs/map/Final_ncfile"
summary(df_Tg)

Tg_nc <- list(df_to_grid(df_Tg,varnam = "Tg", lonnam = "lon", latnam = "lat"))
names(Tg_nc) <- "Tg"
varams = "Tg"
test <- list(lon,lat,Tg_nc,varams)
names(test) <- c("lon","lat","vars","varams")
write_nc2(test,varnams = "Tg",long_name = "Growth temperature",units = "Degree celcius",
          path = "~/data/nimpl_sofun_inputs/map/Final_ncfile/Tg.nc")


``` 


#7. vpd (description and main calculation were provided)
vpd is vapor-pressure-deficient (kPa). We used the 30 yrs average data (1982-2011), which was originally derived from monthly data from CRU ts 4.01. See formula below (Peng et al. 2020 New Phytologist):

vpd <- 0.611*exp(17.27*(Tg)/((Tg)+237.3))-vap*0.1 

where Tg (in celcius degree) was derived from above, and vap (hPa) was derived from CRU ts 4.01 monthly data. The code below shows the global map of final output.

```{r}
library(raster)
library(ncdf4)
ncfname <- paste ("~/data/Final_ncfile/old_9999/", "vpd", ".nc", sep="")
#ncfname <- paste ("E:/R code/CN MS/PhD/NC files/", "vpd", ".nc", sep="")
dname <- "vpd"

ncin <- nc_open(ncfname)

lon <- ncvar_get(ncin,"lon")
nlon <- dim(lon) 
lat<-ncvar_get(ncin,"lat")
nlat <- dim(lat)
vpd <- ncvar_get(ncin,"vpd")

nc_close(ncin)

pre.vec.long <- as.vector(vpd)
pre.vec.long[pre.vec.long == 9999] <- NA

lonlat <- expand.grid(lon, lat)
vpd_input <- as.data.frame(cbind(lonlat,pre.vec.long))
names(vpd_input) <- c("lon","lat","vpd")

df_vpd <- vpd_input

coordinates(vpd_input) <- ~lon+lat 
gridded(vpd_input) <- TRUE
rvpd <- raster(vpd_input, "vpd") 
plot(rvpd)

#output nc file - In Euler its path is same: "~/data/nimpl_sofun_inputs/map/Final_ncfile"
summary(df_vpd)

vpd_nc <- list(df_to_grid(df_vpd,varnam = "vpd", lonnam = "lon", latnam = "lat"))
names(vpd_nc) <- "vpd"
varams = "vpd"
test <- list(lon,lat,vpd_nc,varams)
names(test) <- c("lon","lat","vars","varams")
write_nc2(test,varnams = "vpd",long_name = "Vapor pressure deficient",units = "kPa",
          path = "~/data/nimpl_sofun_inputs/map/Final_ncfile/vpd.nc")

``` 

#8. LMA (a resample method was pre-processed in advance, see text below)
The processing of LMA map, please see LMA_field.Rmd, which was stored in a same file. It has applied (1) resample (2) applied knn method from predictors of PPFD, Tg and alpha. The below code is old and not used any more.

```{r}
#library(raster)
#library(ncdf4)
#SLA_extracted <- read.csv(file="~/data/TRY_maps/SLA_export.csv") #This dataframe outputted from ArcGIS, including FID, lon, lat, Elevation, SLA (mg/mm)

#SLA_extracted <- read.csv(file="E:/C-N cycling/Carbon allocation/672 sites analysis/TRY Global map/SLA_export.csv") #This dataframe outputted from ArcGIS, including FID, lon, lat, Elevation, SLA (mg/mm)
#SLA_extracted2 <- subset(SLA_extracted,SLA_3km_v1>0) #remove NA or negative values

#SLA_extracted2$LMA <- 1000/SLA_extracted2$SLA_3km_v1 #mg/mm -> g/m2

#SLA_extracted3 <- SLA_extracted2[,c(2,3,6)] #only select lon, lat and LMA (g/m2)

#names(lonlat) <- c("lon","lat") #here lonlat was derived from previous coordinates data, which can be continuously used

#final_lma <-Reduce(function(x,y) merge(x = x, y = y, by = c("lon","lat"),all.x=TRUE),list(lonlat,SLA_extracted3)) #combined with global grids

#coordinates(final_lma) <- ~lon+lat 
#gridded(final_lma) <- TRUE
#rLMA <- raster(final_lma, "LMA") 
#plot(rLMA)

#d <- writeRaster(rLMA, paste0("E:/R code/CN MS/PhD/NC files/","LMA_test.nc"),format = "CDF", overwrite = TRUE, varname = "LMA", varunit = "g/m2", longname = "Leaf-mass-per-area (TRY gloabl traits map)",xname = "lon", yname = "lat")


``` 


#9. c3/c4 percentage map
#please see newly created file in c3c4_field.Rmd, as located in the same subfile.
#the below code is old (by arcgis resampling), and not used anymore!
#Although I checked the results after arcgis and R resampling is exactly the same, but we don't use arcgis now....
```{r}
library(raster)
library(ncdf4)

#Now, input file after resampled from ArcGIS (extract multi values from raster)
c4 <- read.csv(file="~/data/c4_still/original/data/export_c4.csv") #This dataframe outputted from ArcGIS, including lon, lat, Elevation, c4 percentage (%)

#c4 <- read.csv(file="D:/PhD/Project 1/Data/ISLSCP_C4/export_c4.csv") #This dataframe outputted from ArcGIS, including lon, lat, Elevation, c4 percentage (%)
summary(c4)

c4a <- subset(c4,c4_percent>=0) #remove NA or negative values

c4a$c4 <- c4a$c4_percent/100 #convert percentage to fraction (0-1)

c4_final <- c4a[,c(1,2,4)] #only select lon, lat and LMA (g/m2)

names(lonlat) <- c("lon","lat")

c4_output <-Reduce(function(x,y) merge(x = x, y = y, by = c("lon","lat"),all.x=TRUE), 
                        list(lonlat,c4_final)) #combined with global grids

coordinates(c4_output) <- ~lon+lat 
gridded(c4_output) <- TRUE
rC4 <- raster(c4_output, "c4") 
plot(rC4)

#d <- writeRaster(rC4, paste0("E:/R code/CN MS/PhD/NC files/","c4_final.nc"),
#                 format = "CDF", overwrite = TRUE, varname = "c4", 
#                 varunit = "fraction", longname = "fraction of c4 plants", 
#                 xname = "lon", yname = "lat")


``` 

